Supermartingales and Termination Analysis of Statistical PCF

### 2021-01-02 LICS21/SPCF.tex
Redefining \tY as a value.
More on def. of \tY and \tY-redex.

The language SPCF is a simply-typed lambda calculus with sampling of real numbers from the closed interval $[0,1]$ and unbounded scoring, following \cite{MakOPW21}. 
Types and terms are defined as follows, where $r$ is a real number, $x$ is a variable, $f : \mathbb{R}^n \to \mathbb{R}$ is any measurable function, and $\Gamma$ is an environment:
\begin{align*}
  \text{\em types } A, B &::= \textsf{R}  \mid  A \to B \\
  \text{\em values } V &::= \lambda x.M  \mid  \underline{r} \\
  \text{\em terms } M, N &::= V  \mid  x  \mid  M_1 \, M_2  \mid  \underline{f}(M_1,\dots ,M_n)  \mid  \tY M \\
   & \qquad \mid  \tif{M < 0}{N_1}{N_2}  \mid  \tsample  \mid  \tscore(M)
\end{align*}
The typing rules are standard (see \Cref{fig:typing rules}).
Terms are identified up to $\alpha$-equivalence, as usual. 
The set of all terms is denoted $\Lambda$, and the set of closed terms is denoted $\Lambda^0$.

>>>>

The language SPCF is a simply-typed lambda calculus with sampling of real numbers from the closed interval $[0,1]$ and unbounded scoring, following \cite{MakOPW21}. 
Types and terms are defined as follows, where $r$ is a real number, $x$ is a variable, $f : \mathbb{R}^n \to \mathbb{R}$ is any measurable function, and $\Gamma$ is an environment:
\akr{\begin{align*}
  \text{\em types } A, B &::= \textsf{R}  \mid  A \to B \\
  \text{\em values } V &::= \lambda x.M  \mid  \underline{r} \\
  \text{\em terms } M, N &::= V  \mid  x  \mid  M_1 \, M_2  \mid  \underline{f}(M_1,\dots ,M_n)  \mid  \tY \, M \\
   & \qquad \mid  \tif{M < 0}{N_1}{N_2}  \mid  \tsample  \mid  \tscore(M)
\end{align*}}
\lo{Currently $\tY$ is not a legal term. This is unnatural: there is no reason to so restrict $\tY$. Following \citep[p.~376]{DBLP:conf/fsttcs/Sieber90}, we should define $\tY$ as a value (and so a proper term); it follows that $\tY \, M$ is redundant as a clause.}
\lo{N.B. It is also standard to regard variables as values in a CBV language.}

\lo{Problem: $\tY \, \tY$ is then a ``stuck'' term: neither value nor redex. A possible solution is to redefine the $\tY$-redex rule as $\tY \, V \to \lambda z . V \, (\tY \, V) \, z$, but I expect this would affect many proofs in the confluent semantics section, and there may not be time for such a change.}
\begin{align*}
  \text{\em types } A, B &::= \textsf{R}  \mid  A \to B \\
  \text{\em values } V &::= x \mid \lambda x.M  \mid  \underline{r} \mid \tY \\
  \text{\em terms } M, N &::= V  \mid  M_1 \, M_2  \mid  \underline{f}(M_1,\dots ,M_n) \\
   & \qquad \mid  \tif{M < 0}{N_1}{N_2}  \mid  \tsample  \mid  \tscore(M)
\end{align*}
The typing rules are standard (see \Cref{fig:typing rules}).
Terms are identified up to $\alpha$-equivalence, as usual. 
The set of all terms is denoted $\Lambda$, and the set of closed terms is denoted $\Lambda^0$.
